Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-03-30 paper/arm.tex
add .tex to input statements for latexdiff
update macros to distinguish REL and ACQ

We discuss \armeight{} using \emph{external global completion} (\EGC{})
\cite{alglave-git-alternate} %,armed-cats},
\cite[\textsection B2.3.6]{arm-reference-manual}
which is very close to our model.

@@

\begin{example}
  Should not wipe out local state only when there is a release-acquire pair,
  as in the definition of \jjr{}, which allows the following execution.  Even
  though there is a blocker $\DWP{x}{3}$, here we are using $\DWP{x}{0}$ to
  justify the lack of dependence $\DRP{x}{4}\xpo\DWP{w}{4}$.
  \begin{gather*}
    \PW{x}{0}\SEMI
    \IF{x}\THEN\PW{y}{2}\FI\SEMI
    \IF{\PR[\mRA]{z}{}}\THEN
      \PR{x}{s}\SEMI
      \IF{s\;\mathsf{even}}\THEN\PW{w}{4}\FI
    \FI
    \PAR
    \PW{x}{2}\SEMI
    \IF{\PR[\mRA]{y}{}}\THEN
      \PW{x}{3}\SEMI
      \PW[\mRA]{z}{1}
    \FI
    \SEMI
    \PW{x}{\PR{w}{}}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{2a}{\DW{x}{0}}{}
        \event{2b}{\DR{x}{2}}{right=of 2a}
        \event{2c}{\DW{y}{1}}{right=of 2b}
        \raevent{2d}{\DR[\mRA]{z}{1}}{right=of 2c}
        \event{2e}{\DR{x}{4}}{right=of 2d}
        \event{2f}{\DW{w}{4}}{right=of 2e}
        %
        \event{3a}{\DW{x}{2}}{below=of 2a}
        \raevent{3b}{\DR[\mRA]{y}{1}}{right=of 3a}
        \event{3c}{\DW{x}{3}}{right=of 3b}
        \raevent{3d}{\DW[\mRA]{z}{1}}{right=of 3c}
        \event{3e}{\DR{w}{4}}{right=of 3d}
        \event{3f}{\DW{x}{4}}{right=of 3e}
        \wk{2a}{2b}
        \po{2b}{2c}
        \sync{2d}{2e}
        \sync[out=20,in=160]{2d}{2f}
        \sync{3b}{3c}
        \sync{3c}{3d}
        \po{3e}{3f}
        \rf{3a}{2b}
        \rf{2c}{3b}
        \rf{3d}{2d}
        \rf{3f}{2e}
        \rf{2f}{3e}
        \wk{2a}{3a}
        %\wk[out=-20,in=-160]{3c}{3f}
        %\wk[out=20,in=160]{2b}{2e}
      \end{tikzinline}}
  \end{gather*}
  (The read-clobbering semantics from \textsection\ref{sec:substitutions}
  does not change this.)
\end{example}

@@

\begin{example}
  \label{ex:dgr1}
  The following example is from \citeauthor{anton} \cite{anton}.  It is
  allowed by \armeight{}, but disallowed by \refdef{def:pomsets-ra}.  The
  coherence order between the writes can be witnessed by a separate thread,
  which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

\begin{example}
  \label{ex:dgr2}
  Revisiting \refex{ex:dgr1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xo{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mRA]{x}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mRA]{x}{2}}{}
          % \xform{x1d}{2{=}r\limplies\bForm}{below=of b}
          \xform{x1i}{\Dx{x} \land\bForm[\FALSE/\Qr{x}]}{below=of b} %x1d}
          % \xo{b}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\DW{y}{1}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Associating right:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{right=.7em of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xos{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mRA]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mRA]{x}{2}}{}
          \event{c}{\Dx{x}\mid\DW{y}{1}}{right=.7em of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing, we have, as desired:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \SEMI
      \PR[\mRA]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
          \event{c}{\DW{y}{1}}{right=of b}
          \rf{a}{b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
\end{example}

% The following execution is allowed both by \armeight{} and \tso{}, but
% disallowed by \refdef{def:pomsets-ra}.
% \begin{gather*}
%   \PW{x}{2}\SEMI 
%   \PR[\mRA]{x}{r}\SEMI
%   \PR{y}{s} \PAR
%   \PW{y}{2}\SEMI
%   \PW[\mRA]{x}{1}
%   \\
%   \hbox{\begin{tikzinline}[node distance=1.5em]
%     \event{a}{\DW{x}{2}}{}
%     \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
%     \event{c}{\DR{y}{0}}{right=of b}
%     \event{d}{\DW{y}{2}}{right=2.5em of c}
%     \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
%     \rf{a}{b}
%     \sync{b}{c}
%     \wk{c}{d}
%     \sync{d}{e}
%     \wk[out=-165,in=-15]{e}{a}
%   \end{tikzinline}}
% \end{gather*}

@@

\begin{example}
  \label{ex:rrd1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  
  \begin{gather*}
    \begin{gathered}
      x\GETS0\SEMI x\GETS 1\SEMI y^\mRA\GETS1
      \PAR
      r\GETS y\SEMI \IF{r}\THEN s\GETS x\FI
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=1.5em]
          \event{wx0}{\DW{x}{0}}{}
          \event{wx1}{\DW{x}{1}}{right=of wx0}
          \raevent{wy1}{\DW[\mRA]{y}{1}}{right=of wx1}
          \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
          \event{rx0}{\DR{x}{0}}{right=of ry1}
          \sync{wx1}{wy1}
          \po{ry1}{rx0}
          \rf{wy1}{ry1}
          \wk[in=-15,out=-165]{rx0}{wx1}
          \wk{wx0}{wx1}
        \end{tikzinline}}
    \end{gathered}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
  dependencies between reads in the locally-ordered-before relation.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wx0}{\DW{x}{0}}{}
        \event{wx1}{\DW{x}{1}}{right=of wx0}
        \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
        \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
        \event{rx0}{\DR{x}{0}}{right=of ry1}
        \sync{wx1}{wy1}
        % \po{ry1}{rx0}
        \rf{wy1}{ry1}
        \wk[in=-15,out=-165]{rx0}{wx1}
        \wk{wx0}{wx1}
      \end{tikzinline}}  
  \end{gather*}
\end{example}

@@

\begin{example}
  Revisiting \refex{ex:rrd1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \xform{x1i}{(1{=}r\lor\RW)\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{r\mid\DR{x}{0}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing sequentially:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\RW)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  At top-level, $\ref{sTOP}$ yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR{y}{r}
      % \SEMI
      % \IF{r}\THEN s\GETS x\FI
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\FALSE)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  The precondition of $\DRP{x}{0}$ is a tautology, as required.
\end{example}

@@

\begin{example}
  \xRRD{} does not adversely affect subscription (\refex{ex:subscription}).
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mRA]{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mRA]{y}{1}}{}
          \xform{x1i}{\Dx{y}\land(\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\mRA}])}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\QL{y}{\mRA}\mid\DW{x}{0}}{}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Since $\QL{y}{\mRA}[\FALSE/\QL{\aLoc}{\mRA}]$ is false, we have:
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mRA]{y}{r}
      \SEMI
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mRA]{y}{1}}{}
          \event{c}{\Dx{y}\land(\RW\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Applying $\ref{sTOP}$ yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR[\mRA]{y}{r}
      % \SEMI
      % \PR{x}{s}
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\TRUE\mid\DR[\mRA]{y}{1}}{}
          \event{c}{\Dx{y}\land(\FALSE\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Although $(\FALSE\limplies\FALSE)$ is a tautology, $\Dx{y}$ is not.  Thus,
  the pomset can only contribute to a top-level pomset if the thread is
  preceded by a relaxed write to $y$, allowing $\DRP[\mRA]{y}{1}$ to be
  downgraded to a relaxed read.
\end{example}

>>>>

We discuss \armeight{} using \emph{external completion} (\EC{})
\cite{alglave-git-alternate} %,armed-cats},
\cite[\textsection B2.3.6]{arm-reference-manual}
which is very close to our model.

@@

\begin{example}
  Should not wipe out local state only when there is a release-acquire pair,
  as in the definition of \jjr{}, which allows the following execution.  Even
  though there is a blocker $\DWP{x}{3}$, here we are using $\DWP{x}{0}$ to
  justify the lack of dependence $\DRP{x}{4}\xpo\DWP{w}{4}$.
  \begin{gather*}
    \PW{x}{0}\SEMI
    \IF{x}\THEN\PW{y}{2}\FI\SEMI
    \IF{\PR[\mACQ]{z}{}}\THEN
      \PR{x}{s}\SEMI
      \IF{s\;\mathsf{even}}\THEN\PW{w}{4}\FI
    \FI
    \PAR
    \PW{x}{2}\SEMI
    \IF{\PR[\mACQ]{y}{}}\THEN
      \PW{x}{3}\SEMI
      \PW[\mREL]{z}{1}
    \FI
    \SEMI
    \PW{x}{\PR{w}{}}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{2a}{\DW{x}{0}}{}
        \event{2b}{\DR{x}{2}}{right=of 2a}
        \event{2c}{\DW{y}{1}}{right=of 2b}
        \raevent{2d}{\DR[\mACQ]{z}{1}}{right=of 2c}
        \event{2e}{\DR{x}{4}}{right=of 2d}
        \event{2f}{\DW{w}{4}}{right=of 2e}
        %
        \event{3a}{\DW{x}{2}}{below=of 2a}
        \raevent{3b}{\DR[\mACQ]{y}{1}}{right=of 3a}
        \event{3c}{\DW{x}{3}}{right=of 3b}
        \raevent{3d}{\DW[\mREL]{z}{1}}{right=of 3c}
        \event{3e}{\DR{w}{4}}{right=of 3d}
        \event{3f}{\DW{x}{4}}{right=of 3e}
        \wk{2a}{2b}
        \po{2b}{2c}
        \sync{2d}{2e}
        \sync[out=20,in=160]{2d}{2f}
        \sync{3b}{3c}
        \sync{3c}{3d}
        \po{3e}{3f}
        \rf{3a}{2b}
        \rf{2c}{3b}
        \rf{3d}{2d}
        \rf{3f}{2e}
        \rf{2f}{3e}
        \wk{2a}{3a}
        %\wk[out=-20,in=-160]{3c}{3f}
        %\wk[out=20,in=160]{2b}{2e}
      \end{tikzinline}}
  \end{gather*}
  (The read-clobbering semantics from \textsection\ref{sec:substitutions}
  does not change this.)
\end{example}

@@

\begin{example}
  \label{ex:dgr1}
  The following example is from \citeauthor{anton} \cite{anton}.  It is
  allowed by \armeight{}, but disallowed by \refdef{def:pomsets-ra}.  The
  coherence order between the writes can be witnessed by a separate thread,
  which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mACQ]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mREL]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mREL]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mACQ]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mACQ]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mREL]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

\begin{example}
  \label{ex:dgr2}
  Revisiting \refex{ex:dgr1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xo{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mACQ]{x}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mACQ]{x}{2}}{}
          % \xform{x1d}{2{=}r\limplies\bForm}{below=of b}
          \xform{x1i}{\Dx{x} \land\bForm[\FALSE/\Qr{x}]}{below=of b} %x1d}
          % \xo{b}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\DW{y}{1}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Associating right:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{right=.7em of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xos{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mACQ]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mACQ]{x}{2}}{}
          \event{c}{\Dx{x}\mid\DW{y}{1}}{right=.7em of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing, we have, as desired:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \SEMI
      \PR[\mACQ]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
          \event{c}{\DW{y}{1}}{right=of b}
          \rf{a}{b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
\end{example}

% The following execution is allowed both by \armeight{} and \tso{}, but
% disallowed by \refdef{def:pomsets-ra}.
% \begin{gather*}
%   \PW{x}{2}\SEMI 
%   \PR[\mACQ]{x}{r}\SEMI
%   \PR{y}{s} \PAR
%   \PW{y}{2}\SEMI
%   \PW[\mREL]{x}{1}
%   \\
%   \hbox{\begin{tikzinline}[node distance=1.5em]
%     \event{a}{\DW{x}{2}}{}
%     \raevent{b}{\DR[\mACQ]{x}{2}}{right=of a}
%     \event{c}{\DR{y}{0}}{right=of b}
%     \event{d}{\DW{y}{2}}{right=2.5em of c}
%     \raevent{e}{\DW[\mREL]{x}{1}}{right=of d}
%     \rf{a}{b}
%     \sync{b}{c}
%     \wk{c}{d}
%     \sync{d}{e}
%     \wk[out=-165,in=-15]{e}{a}
%   \end{tikzinline}}
% \end{gather*}

@@

\begin{example}
  \label{ex:rrd1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  
  \begin{gather*}
    \begin{gathered}
      \PW{x}{0}\SEMI \PW{x}{1}\SEMI \PW[\mREL]{y}{1}
      \PAR
      \PR{y}{r}\SEMI \IF{r}\THEN \PR{x}{s}\FI
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=1.5em]
          \event{wx0}{\DW{x}{0}}{}
          \event{wx1}{\DW{x}{1}}{right=of wx0}
          \raevent{wy1}{\DW[\mREL]{y}{1}}{right=of wx1}
          \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
          \event{rx0}{\DR{x}{0}}{right=of ry1}
          \sync{wx1}{wy1}
          \po{ry1}{rx0}
          \rf{wy1}{ry1}
          \wk[in=-15,out=-165]{rx0}{wx1}
          \wk{wx0}{wx1}
        \end{tikzinline}}
    \end{gathered}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
  dependencies between reads in the locally-ordered-before relation.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wx0}{\DW{x}{0}}{}
        \event{wx1}{\DW{x}{1}}{right=of wx0}
        \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
        \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
        \event{rx0}{\DR{x}{0}}{right=of ry1}
        \sync{wx1}{wy1}
        % \po{ry1}{rx0}
        \rf{wy1}{ry1}
        \wk[in=-15,out=-165]{rx0}{wx1}
        \wk{wx0}{wx1}
      \end{tikzinline}}  
  \end{gather*}
\end{example}

@@

\begin{example}
  Revisiting \refex{ex:rrd1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \xform{x1i}{(1{=}r\lor\RW)\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \IF{r}\THEN \PR{x}{s}\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{r\mid\DR{x}{0}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing sequentially:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN \PR{x}{s}\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\RW)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  At top-level, $\ref{sTOP}$ yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR{y}{r}
      % \SEMI
      % \IF{r}\THEN \PR{x}{s}\FI
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\FALSE)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  The precondition of $\DRP{x}{0}$ is a tautology, as required.
\end{example}

@@

\begin{example}
  \xRRD{} does not adversely affect subscription (\refex{ex:subscription}).
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mACQ]{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mACQ]{y}{1}}{}
          \xform{x1i}{\Dx{y}\land(\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\mRA}])}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\QL{y}{\mRA}\mid\DW{x}{0}}{}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Since $\QL{y}{\mRA}[\FALSE/\QL{\aLoc}{\mRA}]$ is false, we have:
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mACQ]{y}{r}
      \SEMI
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mACQ]{y}{1}}{}
          \event{c}{\Dx{y}\land(\RW\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Applying $\ref{sTOP}$ yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR[\mACQ]{y}{r}
      % \SEMI
      % \PR{x}{s}
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\TRUE\mid\DR[\mACQ]{y}{1}}{}
          \event{c}{\Dx{y}\land(\FALSE\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Although $(\FALSE\limplies\FALSE)$ is a tautology, $\Dx{y}$ is not.  Thus,
  the pomset can only contribute to a top-level pomset if the thread is
  preceded by a relaxed write to $y$, allowing $\DRP[\mACQ]{y}{1}$ to be
  downgraded to a relaxed read.
\end{example}
